\begin{thebibliography}{10}
\footnotesize
\bibitem{Mo94}[Oedra 1994]
  M. odedra-straub
  \newblock The Myths and Illusions of Technology Transfer
  \newblock \emph{IFIP World Congress Proceedings}, 1994.

\bibitem{Jol77}[Jolly 1977]
  Jolly, James A.
  \newblock The Technology Transfer Process: Concepts, Framework and Methodology.
  \newblock \emph{The Journal of Technology Transfer. Springer}, 1977

\bibitem{LV98}[Vaccarezza 1998]
  L. Vaccarezza
  \newblock Ciencia, Tecnología y Sociedad: el estado de la cuestión en América Latina.
  \newblock \emph{Revista Iberoamericana de Educaci\'on, No 18}, 1998


\bibitem{MO90}[Odedra 1990]
  M. Odedra
  \newblock Information {T}echnology {T}ransfer to {D}eveloping
                   {C}ountries: {C}ase studies from {K}enya, {Z}ambia and
                   {Z}imbabwe
  \newblock \emph{PhD thesis London School of Economics}, 1990

\bibitem{R07}[Tovar 2007]
  M. Tovar and R. Rodríguez.
  \newblock PROSPECTIVA Y VIGILANCIA TECNOLOGICA DE LA ELECTRONICA EN COLOMBIA. 
  \newblock \emph{Master thesis, Universidad Nacional de Colombia}, 2007

\bibitem{R99}[Duque 1999]
  M. Duque and A. Gauthier
  \newblock Formación de Ingenieros para la Innovación y el Desarrollo Tecnológico en Colombia
  \newblock \emph{Revista de la Facultad de Minas - Universidad Nacional de Colombia}, 1999


\bibitem{R90}[Zuluaga 2007]
  D Zuluaga, S Campos
  \newblock Informe de Vigilancia Tecnológica: Aplicaciones de la Electrónica en el Sector Agrícola
  \newblock \emph{Technical report, COLCIENCIAS}, 2007

\bibitem{R04}[Martínez 2004]
  H. Martínez
  \newblock Apropiación de conocimiento en Colombia. El caso de los contratos de importación de tecnología
  \newblock \emph{Revista Cuadernos de Economía}, 2004


\bibitem{R02}[Forero 2002]
  C. Forero and H. Jaramillo
  \newblock The access of researchers from developing countries to international science and technology
  \newblock \emph{International Social Science Journal, Volume 54, Issue 171}, 2002

\bibitem{CMS06}[Ostrom 2006]
  C. Hess and E. Ostrom
  \newblock Understanding {K}nowledge as a {C}ommons: {F}rom {T}heory to {P}ractice
  \newblock \emph{Free/Open-Source Software as a Framework for Establishing Commons in Science, MIT Press}, 2006

\bibitem{cohen1}[Cohen 2004]
  Goel Cohen
  \newblock Technology transfer: strategic management in developing countries. 
  \newblock \emph{Publications inc}, 2004

\bibitem{cohen2}[Al-Mabrouk 2008]
  K. Al-Mabrouk and J. Soar. 
  \newblock Building a Framework for Understanding and Improving Information Technology Transfer Process in the Arab Countries.
  \newblock \emph{ 9th IBIMA Conference: Information Management in Modern Organisations - Trends \& Challenges}, 2008

\bibitem{cohen3}[Wood 2009]
  D. Wood and A. Weigel.
  \newblock  International Collaboration on Satellite-Enabled Projects in Developing Countries.
  \newblock \emph{CP1103, Space, Propulsion \& Energy Sciences International Forum - SPESIF}, 2009.

\bibitem{cohen4}[Al-Mabrouk 2009]
  K. Al-Mabrouk and J. Soar.
  \newblock  Identification of key issues for successful technology transfer in the Arab countries: a Delphi study.
  \newblock \emph{International Journal of Technology Transfer and Commercialisation}, 2009.

\bibitem{cohen5}[Janssen 2010]
  R. Janssen. 
  \newblock Exploring the impact of culture Technology transfer to five African countries. 
  \newblock \emph{Master?s thesis, University of Twente, the Netherlands}, 2010.


\end{thebibliography}








